---
title: "Proof of Concept"
bibliography: references.bib
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 2
    self-contained: true
---


```{r suppress_startup_msg, include = FALSE}
# Invisible code chunk to suppress package startup messages later on
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("bamlss"))
suppressPackageStartupMessages(library("digest"))
suppressPackageStartupMessages(library("lemon"))
# Disable readr from showing output when calling read_csv
options(readr.show_progress = FALSE)
options(readr.show_col_types = FALSE)
```


This section/article shows a 'proof of concept' on how to use the
post-processing benchmark data set for post-processing near surface temperature.
While many different post-processing methods have been proposed in
the literature over the past decades, this section makes use of one of the
models proposed by @lang2020, a station-based non-homogeneous Gaussian model
with smoothly varying seasonal effects.





#### Data

@tbl-station shows the three stations used which are chosen to cover different
climatic zones with Emden on the coast of the North Sea, Wasserkuppe in
moderately complex topography in the center of the country, and Oberstdorf
located in a complex topographic environment in the Allgäu Alps at the southern
boarder of Germany.

```{r stations, echo = FALSE}
#| label: tbl-stations
#| tbl-cap: "Stations used in the case study."
knitr::kable(transform(read_csv("eupp_station_meta.csv"),
             altitude = as.integer(round(altitude)), land_usage = NULL))
```

```{r}
#| echo: false
#| fig: true
suppressPackageStartupMessages(library("sf"))
suppressPackageStartupMessages(library("rnaturalearth"))

DEU <- ne_countries(scale = "medium", returnclass = "sf")
DEU <- st_geometry(subset(DEU, adm0_a3_is == "DEU"))
meta <- st_as_sf(read_csv("eupp_station_meta.csv"),
                 coords = c("longitude", "latitude"), crs = st_crs(DEU))

ggplot() + geom_sf(data = DEU) +
    geom_sf(aes(col = station_name), data = meta, cex = 4) +
    ggtitle("Just for Reto, can be removed later")
```

The data set to train the statistical post-processing models consists of
observed near surface temperature ($t_{2m}$) at each of these sites as well
as bi-linearely interpolated forecasts of the near-surface temperature ($ens$)
from the ECMWF ensemble with $50+1$ ensemble members ($0000$ UTC run) covering
the two years 2017 and 2018.

For simplicity, the data set is further restricted to two forecast steps:
$+108h$ and $+120h$ which correspond to  $+4$ days and $12$ hours ($1200$ UTC; noon) and 
$+5$ days ($0000$ UTC; midnight).

For each station, forecast step and date, the ensemble mean
($\overline{\mathrm{ens}}$) as well as the ensemble log-standard deviation
($log(\mathrm{sd}(ens))$) are calculated.


A Python script for downloading
and preparing the data set as described can be
`r xfun::embed_file("etl.py", name = "etl.py", text = "downloaded here")`.






### Methodology

$$
t_{2m} \sim \mathcal{N}(\mu, \sigma),
$$

with

$$
\mu = \beta_0(\mathrm{yday}) + \beta_1(\mathrm{yday}) \cdot \overline{\mathrm{ens}}
$$

$$
log(\sigma) = \gamma_0(\mathrm{yday}) + \gamma_1(\mathrm{yday}) \cdot log(\mathrm{sd}(\mathrm{ens}))
$$






### Estimates and Results

```{r setup}
library("tidyverse")
library("bamlss")
set.seed(6020) # seed for reproducibility
```

```{r utils}
#| code-fold: true
train_models <- function(df, verbose = FALSE) {
    df$log_ens_sd <- log(df$ens_sd)

    f <- list(
        t2m_obs ~ s(yday, bs = "cc") + s(yday, bs = "cc", by = ens_mean),
        ~ s(yday, bs = "cc") + s(yday, bs = "cc", by = log_ens_sd)
    )
    # Caching estimated models for development purposes
    require("digest")
    cached_mod <- sprintf("_cached_bamlss_%s_%s.rds", digest(df), digest(f))
    if (file.exists(cached_mod)) {
        return(readRDS(cached_mod))
    }

    b <- bamlss(f, data = df, verbose = verbose, n.iter = 12000, burnin = 2000, thin = 10)

    saveRDS(b, cached_mod)
    return(b)
}

c95 <- function (x) {
    qx <- quantile(x, probs = c(0.025, 0.50, 0.975), na.rm = TRUE)
    qx <- c(qx[[1]], mean(x), qx[[3]])
    names(qx) <- c("lower", "mid", "upper")
    return(qx)
}

predict_effects <- function(b, file_name) {
    # Extracting station ID and forecast step (in hours) form csv file name
    station <- regmatches(file_name, regexpr("(?<=[_])[0-9]+(?=_)", file_name, perl = TRUE)) # char!
    step    <- as.integer(regmatches(file_name, regexpr("(?<=_)[0-9]+(?=\\.csv$)", file_name, perl = TRUE)))
    
    # Setting up 'grid' (df) with effects to be computed
    args <- data.frame(term  = c("s(yday)", "s(yday,by=ens_mean)", "s(yday)", "s(yday,by=log_ens_sd)"),
                       param = rep(c("mu", "sigma"), each = 2),
                       coef  = paste("varying", rep(c("intercept", "coef"), times = 2), sep = "_"))
            
    # Compute effects
    res <- list()
    nd  <- data.frame(yday = 0:366, ens_mean = 1, log_ens_sd = 1)
    for (i in seq_len(NROW(args))) {
        res[[i]] <- predict(b, nd, term = args$term[i], intercept = FALSE, type = "link", FUN = c95, model = args$param[i]) %>%
                    as_tibble() %>%
                    mutate(yday = nd$yday) %>%
                    pivot_longer(-yday, names_to = "prob") %>%
                    mutate(param = args$param[i], coef = args$coef[i], step = step, station = station)
    }
    return(bind_rows(res))
}
```

```{r read-data}
# Finding and reading available datasets
files <- list.files(pattern = "euppens.*csv")
data  <- setNames(map(files, read_csv), files)
```

```{r model}
# Training post-processing models
models <- setNames(map(data, train_models), files)
```


```{r predict}
# Calculating estimated effects
effects <- map2_dfr(models, files, predict_effects)
```

```{r plot}
#| code-fold: true
d_plt <- effects %>%
    filter(coef == "varying_coef") %>%
    select(-coef) %>%
    mutate(lty = ifelse(prob == "mid", "1", "4"))

gm <- ggplot(subset(d_plt, param == "mu") %>% mutate(param = "beta_1")) +
      geom_line(aes(x = yday, y = value, linetype = lty, group = interaction(station, prob), col = station)) +                         
      facet_grid(rows = vars(step), cols = vars(param)) +
      ylab("") +
      guides(linetype = "none") +
      theme_minimal()

gs <- ggplot(subset(d_plt, param == "sigma") %>% mutate(param = "gamma_1")) +
      geom_line(aes(x = yday, y = value, linetype = lty, group = interaction(station, prob), col = station)) +                         
      facet_grid(rows = vars(step), cols = vars(param)) +
      ylab("") +
      guides(linetype = "none") +
      theme_minimal()


library("lemon")
grid_arrange_shared_legend(gm, gs, ncol = 2, position = "right")
```

```{r plot2}
#| code-fold: true
d_plt_ic <- effects %>%
    filter(coef == "varying_intercept") %>%
    select(-coef) %>%
    mutate(lty = ifelse(prob == "mid", "1", "4"))

gm_ic <- ggplot(subset(d_plt_ic, param == "mu") %>% mutate(param = "beta_0")) +
         geom_line(aes(x = yday, y = value, linetype = lty, group = interaction(station, prob), col = station)) +                         
         facet_grid(rows = vars(step), cols = vars(param)) +
         ylab("") +
         guides(linetype = "none") +
         theme_minimal()

gs_ic <- ggplot(subset(d_plt_ic, param == "sigma") %>% mutate(param = "gamma_0")) +
         geom_line(aes(x = yday, y = value, linetype = lty, group = interaction(station, prob), col = station)) +                         
         facet_grid(rows = vars(step), cols = vars(param)) +
         ylab("") +
         guides(linetype = "none") +
         theme_minimal()

library("lemon")
grid_arrange_shared_legend(gm_ic, gs_ic, ncol = 2, position = "right")
```

```{r combine-plots}
grid_arrange_shared_legend(gm_ic, gm, gs_ic, gs, ncol = 4, position = "bottom")
```


